---
title: "Readings lecture 1 - MDS"
author: "Bo Crauwels"
date: "1/24/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# R Tutorial Intro

## Basic data types 

Numeric data: 
```{r}
x = 10.5
class(x)
```

```{r}
k = 10
class(k)
is.integer(k)
```

Integer data: 
```{r}
y = as.integer(3)
class(y)
is.integer(y)
as.integer(3.141592) #coerce decimal to integer 
as.integer("3.141592") #coerce string to integer
as.integer('Bo') #cannot coerce non-decimal string to integer
```

Complex value: 
```{r}
z = 1 + 2i #complex values are defined by i (imaginary value)
class(z)
```

Logical value: 
i.e., pften a comparison between variables
```{r}
x = 1; y=2 #creating two variables in same line 
z = x>y
z
class(z)
```

```{r}
u = TRUE; v = FALSE# logical operations are &, | and !
u&v
u|v
!u
```

Character values: 
```{r}
x = as.character(32.3)
class(x)
fname = "bo"; lname = "crauwels"
paste (fname, lname)
sprintf("%s has %d dollars", "Bo", 20)#readable string w sprintf function 
sub("little", "big", "Bo has a little room")
```

##Vector 

A vector is a sequence of data elements of the same basic type. 
```{r}
c(2,3,5)
c(T,F,T,F)
c("a", "b", "c")
length(c("a", "b", "c"))
```

Combining vectors: 
```{r}
n = c(2,3,5); s = c("a", "b")
c(n,s)
```

Vector arithmetics: 
```{r}
a = c(1,3,5,6)
b = c(1,2,3,4)
5*a
a+b
a = c(a,2)
a+b #recycling rule, b will be recycled cause shorter
```

Vector index: 
```{r}
s = c("aa", "bb", "cc", "dd")
s[4]
s[-1]
```

Numeric index vector: 
```{r}
s = c("aa", "bb", "cc", "dd", "ee")
s[c(2,3)]
s[c(2:4)]
```

Logical index vector: 
```{r}
s = c("aa", "bb", "cc", "dd", "ee")
L = c(F,T,T,F,F)
s[L]
```

Named vector members: 
```{r}
v = c("Mary", "Sue")
names(v)= c("first", "last")
v
v[c("last", "first")]
```

##Matrix 

Matrices are data arranged in a two-dimensional rectangular layout. Creating a matrix: 
```{r}
A = matrix(
  c(2,4,3,1,5,7), #data input
  nrow = 2,
  ncol=3,
  byrow = T
)
A
A[2,3]
A[2,]
A[,3]
A[, c(2,3)]

dimnames(A)= list(
  c("row1", "row2"),
  c("col1", "col2", "col3")
)
A
```

Matrix construction: 
```{r}
t(A) #transpose

B = matrix(
  c(1,1,1,1,1,1), #data input
  nrow = 2,
  ncol=3
)

cbind(A,B) #column bind
rbind(A,B) #row bind 
c(B)#deconstruct matrix 
```

##List 

A list is a vector containing other objects
```{r}
n = c(2,3,5)
s = c("aa", "bb", "cc", "dd")
b = c(T,T,F)
x = list(n,s,b,3)
x #contains copies of above objects
x[2] #retrieving list member
x[c(2,3)]
x[[2]] #referencing list member

# a referenced list member is not the original member, so you can modify it without modifying the original member 

x[[2]][1] = "hey"
x[[2]]
s
```

Named list members:
```{r}
v = list(bob=c(2,3), john = c(1,2))
v['bob']
v[c("john", "bob")]
```

##Data frame: 

Data frames are used for storing data tables. It's a list of vectors of equal length. 

```{r}
n = c(2,3,5)
s = c("aa", "bb", "cc")
b = c(T,T,F)
df = data.frame(n,s,b)
nrow(df)
ncol(df)
head(df)
```

df column vector: 
```{r}
df[[2]]
```

df column slice: 
```{r}
df[2]
df[c(1,2)]
```

df row slice: 
```{r}
df[1,]
df[c(1,2),]
```

logical indexing: 
```{r}
L = df == 3
L
df[L,] #list of rows in df that contain 3 
```


